\chapter*{Acknowledgments}
\label{Acknowledgments}
\markboth{}{}
%Looking back at the journey to pursue my Ph.D., I realize how many things happened in just four years. This thesis would probably have been never finished without the support of many people, who I would like to thank here.  
I would like to thank here all the people who contributed to this work and supported me during the journey to pursue my Ph.D.\\\\
First and foremost, my gratitude goes to Dr. Wolfgang Kelsch for supervising me, and for the interesting discussion and meeting, which contributed to both this work and my professional growth.\\\\
Thanks to Dr. Eleonora Russo for supervising me on the statistical analysis, and for the fruitful discussions on this project.\\\\
Thanks to Max Scheller for providing me the data and the unit classification, and for the vivid discussion during our weekly meetings together with Dr. Wolfgang Kelsch and Dr. Eleonora Russo.\\\\
I would like to extend my appreciation to Dr. Georgia Koppe for providing with her scripts and for the time she dedicated to me with further explanations.\\\\
Thanks to Prof. Dr. Daniel Durtstewitz, thanks to whom my experience at ZI is started.\\\\
I would like to express my gratitude to all members of my doctorate committee, Prof. Dr. Daniel Durstewitz, Prof. Dr. Christoph Schuster, Dr. Kevin Allen for having agreed to referee my thesis and for their fruitful comments.\\\\
Thanks to my current and former officemates: Cathy, David, Laurens, Mirko, Sebastian, Hazem, Georgia and Dominik because it is a pleasure to work in such a friendly environment.\\\\
A special thought goes to Dr. Melanie Fritz for the pleasant and useful conversations.\\\\
From my heart I would like to thank my family. I thank my parents for always supporting and trusting me; there is no distance able to cancel my gratitude. I thank my sister Melania, because she is always able to motivate me, and during these years was never tired to listen me, for any talk rehearsal and request of advice. I would like thank Carlo for the interesting statistical discussion and the useful tips.\\\\
A special thank goes to the grace of my life, Guido, who knows any emotional state of this work and always supported me with empathy and gentleness.
