\chapter*{Acknowledgments}
\markboth{}{}
Looking back at the journey to pursue my PhD, I realize how many things happened in just four years. This thesis would probably have been never finished without the support of many people, who I would like to thank here.  

First and foremost, my gratitude goes to Dr. Wolfgang Kelsch for supervising me, and for any discussion and meeting, which were always interesting and contributed to both this work and my professional growth. 

Thanks to Dr. Eleonora Russo for the supervision on the statistical analysis, for her constant availability and for the fruitful discussions on this project.

Thanks to Max Scheller for providing me the data and for the useful clarifications and discussions both about the experimental dataset and the analysis.


I would like to express my gratitude to all members of my doctorate committee, Prof. Dr. Daniel Durstewitz, Prof. Dr. Christoph Schuster, Dr. Kevin Allen for having agreed to referee my thesis and for their fruitful comments.  

I would extend my appreciation to Georgia for providing with her scripts and for the time she dedicated to me with further explanations.


Thanks to my current and former officemates: Nitin, Thomas, Mohammad, Harshit, Behnaz, Deepesh, Yuriy and Pranav for helping me to remain focused every day.
In particular, thanks to Deepesh and Behnaz for having supported me and for the nice and pleasant time we spent together. I will be always grateful for your friendship. 

I will never forget the kindness and thoughtfulness you had for me.



During these years, I also had the opportunity to join the \emph{Kabelstraat} events. I am thankful to all people there for the funny evenings and good dinners together.

I wish to thanks my friends and former colleagues back in Italy: Federica, Ida, Stefania, Daniela and Arianna. It was nice to study with you at the university and it looks amazing to me that after many years we are still in touch.

Misa, Ale G., Ghislaine, Chiara, Manouk, Tsvetan and Jacopo, I will never regret to have shared a house with you. Thanks for never making me feel lonely coming back home after work. Thanks to you, I discovered the beauty and the difficulty to meet up with people of different cultures.

During these years I found a concentration of Italians at the fourth floor of our department. Thanks Ale G., Ale Z., Carlo, Enrico, Fabio, Gianmarco, Lorenzo, Marta and Misa for making me feel like home. I will miss our barbecue parties in Tongelresestraat.\\

Before starting my PhD, I would never thought of start doing climbing in a flat country. This sport has the power to make my mind rest after long days in the office.
I would like to thank all my climbing friends for motivating me and for sharing with me this passion. Most of all, Ale Z. for being my personal climbing teacher, and Carlo for making me discover how wonderful is to climb on the real rock.\\ 

A special thank goes to my family, especially to my parents and my sister.

Dad, you have been always a mystery for me. I am not able to explain all my emotions, but I can say for sure that the strongest feelings I have for you are love and endearment. Thank you for all you did for our family. I like thinking you have been at my side for all these years, and that now you are proud of me.

Thanks to my mother for her support. Mamma, grazie per avermi supportato in ogni momento e non aver mai ostacolato le mie scelte nonostante a volte abbiano causato sofferenza e nostalgia per la distanza che ci separa. Il tuo amore e la tua fiducia in me, sono stati fondamentali per il raggiungimento di questo traguardo. Ti voglio bene.   

To my sister Carla for having been part of my \textit{entire} life. Sister, I cannot put into words how grateful I am, and how much I owe you. I would have never been able to achieve none of my successes without you.
I miss sharing the daily life with you, but I feel so lucky when I realize that the distance did change nothing. 

I desire to thank Guido for having organized very beautiful excursions around the south of Germany and for the nice time we spent in the Netherlands with Carla and Carlo. I have a good memory of all the trips we made together. You are a good and warm-hearted person.

Last, but not least, I wish to thank you, Carlo, for your love. 
Thanks for bringing happiness, color and brightness to my life. I cannot imagine my future without you.
%\clearpage{\pagestyle{empty}\cleardoublepage}
\addcontentsline{toc}{chapter}{Acknowledgments}